---
title: "linear regression course"
author: "Mohamed Elashri"
date: "May 18, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Task 
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

"Is an automatic or manual transmission better for MPG"

"Quantify the MPG difference between automatic and manual transmissions"



## Data preperation 
```{r}
data(mtcars)
names(mtcars)

```


and this is the summary of dataset information 
```{r}
summary(cars)

```

There are eleven attributes in this dataset, we want to get what is the relation between MPG attribute and the other attributes.



## Which is better ?

To get the relation we should first see the correlation between the MPG and the others, which is the first step. Thanks R that it is very easy to do this using built-in function cor().

```{r}
cor(mtcars$mpg,mtcars[,-1])

```

we see that there are positive correlation whith are (drat, qsec, vs, am, and gear) and negative correlation for the others (cyl, disp, hp, wt and carb). 

Now to answer the transmision part we should first assign values to this catogrical attribute (transmission type) as 
0 -> automatic transmission type
1 -> manual transmission type

doing this with R as 
```{r}
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <-c("Automatic", "Manual")

```

we can see which is better from using boxplot (more about boxplot idea here https://towardsdatascience.com/understanding-boxplots-5e2df7bcbd51 ) and plot is in appendix 1 


The plot tells us the manual have better MPG than the automatic .But we need to test this claim (we want to see if we can reject the null hypothesis). 

perform t-test for example using R 
```{r}
t.test(mtcars$mpg~mtcars$am,conf.level=0.95)

```

We can see p-value of 0.001374 which gives us clue that we should reject the null hypothesis which is that there is not difference in MPG. This mean that the automatic cars has lower mpg than the manual cars.  

The conculsion is based on the assumtion that the automatic and manual cars other 10 characteristics are the same which is not accurate and to invistigate this we should apply multiple regression models to fit more attributes in our analysis (see Appendix 3 for details).

Now I want 

## MPG Difference Qualification

First, we can try to do a multivariate linear regression with all attributes (see appendix 3)
```{r}
model = lm(data = mtcars, mpg ~ .)
summary(model)

```
If we look to the coefficiente we observe that wt is the only attribute changing with our mpg so we don't use all variables to fit our model because it will probably result in overfitting problem. So we need to test different models with different exploratory variables.

We use R function step() which is this automatic model-choosing function that choose the best linear regression model

```{r}
bestmodel = step(lm(data = mtcars, mpg ~ .), trace=0)
summary(bestmodel)

```

 looks like the best model is the one that includes wt, qsec and am, So they also  needs to be considered. wt negatively changes with mpg, and qsec and am positively changes. Every kg/1000 weight increase will cause a decrease of roughly 4 mpg, every increase of 1/4 mile time will cause an increase of 1.2 mpg, and on average, manual transmission is 2.9 mpg better than automatic transmission. The model is able to explain 85% of variance. The residual plots also seems to be randomly scattered (see appendix 4).


##  The colclusion 

Based on the previous analysis, we can say that on average manual transmission is better than automatic transmission by 2.9 mpg but also transmission type is not the only factor accounting for MPG, weight, and acceleration (1/4 mile time) also needs to be considered. 

### Appendix 1 (Barplot and vilion plot)

```{r}
boxplot(mtcars$mpg ~ mtcars$am, data = mtcars, outpch = 19, ylab="MPG",xlab="transmission type",main="mpg vs transmission type", col="red")

```

We can also get same information from vilion plot.

```{r}
library(stats) 
library(ggplot2)
ggplot(mtcars, aes(y=mpg, x=factor(am, labels = c("automatic", "manual")), fill=factor(am)))+
        geom_violin(colour="black", size=1)+
        xlab("transmission") + ylab("MPG")

```

## Appendix 2 (Residual Plots)


```{r}
par(mfrow = c(2,2))
plot(bestmodel)

```

## Appendix 3 (Regression Model Results)


comparison between best model to fit the dataset and simple regression model that we used. 
```{r}
full.model <- lm(mpg ~ ., data = mtcars)
best.model <- step(full.model, direction = "backward")

```



